---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

```{r}

# Libraries
library(tidyverse)
library(psych)
library(usmap)

```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```

```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}

# Descriptive stats from psych package
describe(inc)

# Factor industry
inc$Industry <- factor(inc$Industry)
summary(inc$Industry)

# Factor state
inc$State <- factor(inc$State)
summary(inc$State)

# Bin growth rate
summary((cut(inc$Growth_Rate, breaks=10)))

# Bin revenue
summary((cut(inc$Revenue, breaks=10)))

# Bin employees
summary((cut(inc$Employees, breaks=10)))

```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, fig.width=6, fig.height=8}

# Create frequency table
dfstate <- data.frame(table(inc$State))
colnames(dfstate) <- c('state', 'freq')

# Bar plot
dfstate %>%
    ggplot(aes(x=reorder(state, freq), y=freq)) +
    geom_bar(stat='identity') +
    coord_flip() +
    theme_minimal() +
    ggtitle('Company Counts by State') +
    xlab('') + ylab('Companies')

```

```{r}

# Map plot
plot_usmap(data = dfstate, values='freq') + 
    scale_fill_continuous(name='Companies', label=scales::comma, low='cornsilk', high='brown4') + 
    theme(legend.position='right') +
    ggtitle('Company Counts by State')

```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

Two options are shown. First, an ordinary box plot, which shows the inter-quantile range, median, and outliers, but which doesn't show the distribution of values.


```{r}

# Filter out NY companies, only using complete observations
inc_ny <- inc[complete.cases(inc),]
inc_ny <- inc[inc$State=='NY',]

# Boxplot
inc_ny %>%
    ggplot(aes(x='', y=Employees)) +
    geom_boxplot(width=0.75, color='black', fill='magenta4') +
    stat_summary(fun='mean', geom='point', shape=20, size=5, color='grey', fill='grey') +
    scale_y_continuous(trans='log', breaks=c(10, 100, 1000, 10000)) +
    theme_minimal() +
    ggtitle('New York Company Employee Counts') +
    ylab('Employees (log scale)') + xlab('') +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
    theme(legend.position='right') +
    geom_point(aes(shape='mean'), alpha=0) +
    guides(shape=guide_legend(title=NULL, override.aes=list(alpha=1, color='grey', size=5, shape=20)))

```

The second figure includes a violin plot overlain onto the box plot which shows the distribution. The violin has the disadvantage that the outliers have to be added manually using geom_point().


```{r}

# Outlier detection
out_lo <- quantile(inc_ny$Employees, probs=c(0.25)) - (1.5 * IQR(inc_ny$Employees))
out_hi <- quantile(inc_ny$Employees, probs=c(0.75)) + (1.5 * IQR(inc_ny$Employees))

# Violin plot
inc_ny %>%
    ggplot(aes(x='', y=Employees)) +
    geom_violin(color='black', fill='magenta4', draw_quantiles=c(0.25, 0.5, 0.75)) +
    geom_point(data=inc_ny[inc_ny$Employees > out_hi | inc_ny$Employees < out_lo,]) +
    stat_summary(fun='mean', geom='point', shape=20, size=5, color='grey', fill='grey') +
    scale_y_continuous(trans='log', breaks=c(10, 100, 1000, 10000)) +
    theme_minimal() +
    ggtitle('New York Company Employee Counts') +
    ylab('Employees (log scale)') + xlab('') +
    theme(legend.position='right') +
    geom_point(aes(shape='mean'), alpha=0) +
    guides(shape=guide_legend(title=NULL, override.aes=list(alpha=1, color='grey', size=5, shape=20)))

```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r, fig.height=8, fig.width=6}

# Aggregate revenue by industry and convert to billions
inc_rev <- inc[complete.cases(inc), ] %>%
    group_by(Industry) %>%
    summarise(rev=round(sum(Revenue)/sum(Employees)/1000, 1))

inc_rev %>%
    ggplot(aes(x=reorder(Industry, rev), y=rev))+
    geom_bar(stat='identity', width=0.5) +
    coord_flip() +
    theme_minimal() +
    ggtitle('Revenue per Employee by Industry') +
    xlab('') + ylab('Revenue per Employee (Thousands of USD)')

```


